Clustering in a hyperbolic model of complex networks.

### 2020-03-10 hyperbolic_degrees.tex
Edits by Tobias

\section{Degrees in the KPKVK model: proof of Theorem~\ref{thm:degrees_hyperbolic}}

@@

We start by using the Campbell-Mecke formula to compare the degree distribution in $\GPo$ with that of a \emph{typical point} in $\Ginf$. For this we will be considering the Poisson process $\Pcal$ to which we add one additional point $(0,y)$ on the $y$-axis. In some computations the height $y$ will be fixed, but eventually we shall take it exponentially distributed with parameter $\alpha$, and independent of $\Pcal$.
We refer to $(0,y)$ as ``the typical point''.

To provide some intuition for this definition and name, note that we can alternatively view $\Pcal$ as follows. 
We take a constant intensity Poisson process on $\eR$ corresponding to the $x$-coordinates, and to each point
we attach a random ``mark'', corresponding to the $y$-coordinate, where the marks are i.i.d.~exponentially distributed with parameter $\alpha$. By translation invariance of Poisson point process, each point will behave as $(0,y)$ with $y$ exponentially distributed with parameter $\alpha$, the typical point of the process.

We first compute the degree distribution of the typical point in $\Ginf$ and show that this equals 
\[
	p_k = \int_0^\infty \rho(y,k) \alpha e^{-\alpha y} \, dy.
\]
We then relate this to the Poissonized KPKVB model $\GPo$. More precisely, let $N_{\Po}(k)$ denote the set of degree $k$ vertices in $\GPo$. We then show in Section~\ref{ssec:expected_degrees_GPo} that for any $1 \ll k_n \le n -1$ with $k_n = \bigO{n^{\frac{1}{2\alpha + 1}}}$, $\Exp{N_{\Po}(k_n)} = (1+\smallO{1}) n p_{k_n}$ and more generally,
\begin{equation}\label{eq:def_factorial_moments_GPo}
	\Exp{\binom{N_{\Po}(k_n)}{r}} = (1+\smallO{1}) \frac{\Exp{N_{\Po}(k_n)}^r}{r!},
\end{equation}
for any integer $r \ge 1$ in Section~\ref{ssec:factorial_moments_GPo}. The latter result requires us to analyze the joint degree distribution in $\GPo$, which we do in Section~\ref{ssec:joint_degrees_GPo}. The above result in particular implies concentration of $N_{\Po}(k_n)$ from which the result on the degree distribution in $\GPo$ follows for $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. When $k_n = (1+\smallO{1}) c n^{\frac{1}{2\alpha + 1}}$ we use the above result to show that the fraction of degree $k_n$ nodes in $\GPo$ converges to a Poisson distribution.

@@

We will also establish all the above mention results for the finite box model $\Gbox$, since the proofs only require small alterations and we will need these results later on when analyzing the clustering coefficient and function.

\subsection{The degree of the typical point in $\Ginf$}\label{ssec:degrees_infinite_model}

We start with investigating the distribution of the degree of the typical point. (Considered in the graph obtained from $\Ginf$ by adding the typical point to its vertex set.)

If $p = (x,y) \in \eR\times[0,\infty)$ is a point, not necessarily part of the Poisson process, then we will write
\[ 
	\mu(y) = \mu(p) := \mu(\BallPo{p}).
\]
Integrating the intensity function of $\Pcal$ over $\BallPo{p}$ gives, using $\alpha>\frac{1}{2}$,
\begin{align*}
	\mu(y) &= \int_0^\infty \int_{-e^{(y+y')/2}}^{e^{(y+y')/2}} \frac{\alpha \nu}{\pi} e^{-\alpha y'} \, dx' \, dy' 
		= \frac{2\alpha \nu e^{y/2}}{\pi} \int_0^\infty e^{(\frac12-\alpha)y'} \, dy' 
		= \frac{2\alpha \nu e^{y/2}}{\pi(\alpha-\frac12)} = \xi e^{y/2}.
\end{align*}

For $p = (x,y) \in \eR\times[0,\infty)$ we define
\begin{equation}\label{eq:def_degree_inf_model}
	\rho(p,k) := \Prob{\Po(\Mu{p}) = k},
\end{equation}
where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$. When $p = (0,y)$ we often write $\rho(y,k)$ instead of $\rho(p,k)$.

Let the random variable $D$ denote the degree of the typical point. Since the typical point has a height that is independent of the Poisson process and exponential($\alpha$)-distributed, for $k \in \N_0$:
\begin{equation}\label{eq:def_pk}
	p_k :=  \Prob{D = k} = \int_0^\infty \rho(y,k) \alpha e^{-\alpha y} \, dy.
\end{equation}

Using the transformation of variables $z = \xi e^{\frac{y}{2}}$ (so $dy = \frac{2}{z}dz$), we compute
\begin{align*}
	p_k
    &= \frac{1}{k!} \int_0^\infty \left(\xi e^{\frac{y}{2}}\right)^k 
    	e^{-\xi e^{\frac{y}{2}}} \alpha e^{-\alpha y} \, dy
    	= \frac{\alpha \xi^{2\alpha}}{k!} \int_0^\infty 
    	\left(\xi e^{\frac{y}{2}}\right)^{k - 2\alpha} e^{-\xi e^{\frac{y}{2}}}
        \, dy\\
    &= \frac{2\alpha \xi^{2\alpha}}{k!} \int_{\xi}^{\infty} 
    	z^{k -2\alpha-1} e^{-z} \, dz
    	= \frac{2\alpha \xi^{2\alpha}\Gamma^+(k - 2\alpha, \xi)}{k!},
\end{align*}
where we recall that $\Gamma^{+}$ denotes the upper incomplete Gamma-function.
Note that, unsurprisingly, this is identical to the expression Gugelmann et al.~\cite{gugelmann2012random} gave for the limiting degree distribution of $G(n;\alpha,\nu)$. Using Stirling's approximation to the gamma function, we find that

\begin{equation}\label{eq:degree_distribution_P_asymptotics}
	p_k \sim 2\alpha\xi^{2\alpha} k^{-(2\alpha + 1)}
	\quad \text{as } k \to \infty.
\end{equation}

By a similar computation we have the following result, which will be useful later on. For any $\beta > 0$, as $k \to \infty$
\begin{equation}\label{eq:general_integral_rho_y_k}
	\int_0^\infty e^{\beta y} \rho(y, k) \alpha e^{-\alpha y} \, dy
    \sim 2\alpha \xi^{2(\beta + \alpha)} k^{-2(\beta + \alpha)-1}.
\end{equation}

@@

We can now use a concentration of heights argument to show that the integration of the Poisson probabilities $\Prob{\Po(\mu_{\Po}(y)) = k_n}$ over $0 \le y \le (1-\varepsilon) R$ is asymptotically equivalent to $p_{k_n}$. And the same holds if we instead consider $\mu_{\mathrm{box}}(y)$. The proof contains some technical elements that are contained in the Appendix to not hinder the flow of the argument.

\begin{lemma}\label{lem:degree_integral}
Let $0 < \varepsilon < 1$. Then for all $0 \le k_n \le n - 1$, as $n \to \infty$,
\begin{equation}\label{eq:degree_integral_BallHyp}
	\int_0^{(1-\varepsilon)R} \Prob{\Po(\Mu{\BallHyp{y}} = k_n} \alpha e^{-\alpha y} \dd y
	= (1+\smallO{1}) p_{k_n},
\end{equation}
where
\begin{equation}\label{eq:def_average_degree_Gbox_phi}
	p_{k_n} = \int_0^\infty \Pee(Po(\xi e^{\frac{y}{2}})=k_n)\alpha e^{-\alpha y}dy.
\end{equation}
Moreover, the same holds if we replace $\Mu{\BallHyp{y}}$ with $\Mu{\BallPon{y}}$ in~\eqref{eq:degree_integral_BallHyp}.
\end{lemma}

\begin{proof}
We will show that
\[
	\int_0^{(1-\varepsilon)R} \Prob{\Po(\mu_{\Po}(y)) = k_n} \alpha e^{-\alpha y} \dd y
	= (1+o(1))\int_0^\infty \Pee(\Po(\xi e^{\frac{z}{2}})=k_n) \alpha e^{-\alpha z} \dd z.
\]
This implies the result because the last integral equals $(1+o(1))2\alpha \xi^{2\alpha} k_n^{-(2\alpha+1)} = (1+o(1))p_{k_n}$.

@@

\begin{lemma}[First moment of number of degree $k$ vertices]\label{lem:expnnkn}
Let $N_{Po}(k)$ and $\Nbox(k)$ denote the number of vertices with degree $k$ vertices in the poissonized KPKVB model $\GPo$ and the finite box model $\Gbox$, respectively. Consider a sequence of integers $k_n\rightarrow\infty$ with $0 \leq k_n \leq n-1$.	If $k_n = O\left(n^{\frac{1}{2\alpha+1}}\right)$, then as $n \to \infty$,
\[
	\Exp{N_{Po}(k_n)} =(1+o(1)) n p_{k_n} \quad \text{and} \quad \Exp{\Nbox(k_n)} =(1+o(1)) n p_{k_n}.
\]
Moreover, if $k_n \gg n^{\frac{1}{2\alpha + 1}}$ then
\[
	\Exp{N_{\Po}(k_n)} = \smallO{1} \quad \text{and} \quad \Exp{\Nbox(k_n)} = \smallO{1}
\]
\end{lemma}

@@

When $k_n \gg n^{\frac{1}{2\alpha + 1}}$, Lemma~\ref{lem:degree_integral} implies that
\[
	n \int_0^{(1-\varepsilon)R} \Pee(Po(\mu_{Po}(y))=k_n) \alpha e^{-\alpha y}dy
	= (1+\smallO{1}) n p_{k_n} = \bigO{n k^{-(2\alpha + 1}} = \smallO{1}.
\]
On the other hand,
\[
	n \int_{(1-\varepsilon)R}^R \Pee(Po(\mu_{Po}(y))=k_n) \alpha e^{-\alpha y}dy = \bigO{n^{1 - 2\alpha(1-\varepsilon)}}.
\]
which is $\smallO{1}$ since by our choice $2\alpha(1-\varepsilon) > 1$. Thus the second claim of the lemma follows.
\end{proof}

@@

\begin{lemma}[Asymptotic factorization of degree probabilities]\label{lem:probdegFact}
Let $C, C^\prime>0$ and $r,s$ be positive integers with $r+1\leq s$. Fix $0<\varepsilon<1$. Then, it holds uniformly for all $(v_1,\dots,v_s) \in \left(\stripknc\right)^{s} = \stripknc \times \dots \times \stripknc$, satisfying $|x_{v_i}-x_{v_{r+1}}|\geq k_n^{1+\varepsilon}$ for all  $1\leq i \leq r$,  that
\[
	\varphi(\{v_1,\dots,v_{r+1}\};v_1,\dots,v_s)
	=(1+o(1))\varphi(\{v_1,\dots,v_r\};v_1,\dots,v_s)\varphi(\{v_{r+1}\};v_1,\dots,v_s)+\bigO{k_n^{-C^\prime}},
\]
where $\varphi$ is either $\varphi_{\Po}$ or $\varphi_{\mathrm{box}}$.
(Here uniformity means that the $\smallO{1}$ and $\bigO{k_n}$ terms do not depend on $v_1,\dots,v_{s}$.)
\end{lemma}

@@

We will now use that $\frac{(a+b)!}{a!} = (1+o(1))(a+b)^b$ for $b^2=o(a)$, applied to $a=k_n-y_{r+1}$ and $b=y_{r+1}-s$. To see this auxiliary fact, note that by Stirling's approximation to the factorial, it follows that

@@

\begin{lemma}[Factorial moments]\label{lem:factmoment}
Recall that $N_{Po}(k)$ denotes the number of degree $k$ vertices in $\GPo$. Then, for any positive integer $r$ and $k_n \rightarrow \infty$, it holds that
\[
	\Exp{{N_{Po}(k_n) \choose r}} =(1+o(1)) \frac{(\Exp{ N_{Po}(k_n)})^r}{r!}.
\]
\end{lemma}

@@

\begin{lemma}[Factorization of degrees]
	\label{lem:asympind}
	Let $\varphi$ be either $\varphi_{\mathrm{box}}$ or $\varphi_{\Po}$. Then we have that
	\begin{align*}
	&\int_\Rcal \cdots \int_\Rcal \varphi(\{p_1,\dots,p_r \}, k_n ;p_1,\dots,p_r) 
		f(x_1,y_1) \dd x_1 \dd y_1 \cdots f(x_r,y_r)\dd x_r \dd y_r \\
	&= (1+o(1))\left(\int_\Rcal  \varphi(\{p_1\}, k_n ;p_1) f(x,y) \dd x \dd y\right)^r .
	\end{align*}
\end{lemma}

@@

First of all, we observe that
\begin{align*}
{N_{\Po}(k_n) \choose r} = \frac{1}{r!} \sum_{p_1,\dots,p_r\in V(\GPo)}^{\not = } \1_{\{D_{\GPo}(p_1)=\dots=D_{\GPo}(p_r)=k_n\}}.
\end{align*}
This can be seen by induction on $r$. For $r=1$, the claim is clear. Assuming it holds for $r\geq 1$, by the induction hypothesis,
\begin{align*}
&{N_{\Po}(k_n) \choose r+1} = {N_{\Po}(k_n) \choose r} \frac{N_{\Po}(k_n)-r}{r+1} \\
&= \frac{1}{(r+1)!}\sum_{p_1,\dots,p_r\in V(\GPo)}^{\not = } \1_{\{D_{\GPo}(p_1)=\dots=D_{\GPo}(p_r)=k_n\}} (N_{\Po}(k_n)-r).
\end{align*}
Now, we can write
\[
	N_{\Po}(k_n) = \sum_{\stackrel{p_{r+1} \in V(\GPo),}{p_{r+1} \not \in\{ p_1,\dots,p_r\}} } \1_{\{D_{\GPo}(p_{r+1})=k_n\}}+\sum_{p_{r+1} \in \{p_1,\dots,p_r\}}\1_{\{D_{\GPo}(p_{r+1})=k_n\}}.
\]
The first sum leads to the right-hand side of the claim for $r+1$, whereas the second sum will cancel with the $-r$.

By the Campbell-Mecke formula 
\begin{align*}
	&\Exp{ {N_{\Po}(k_n) \choose r}}=\frac{1}{r!} \E\left[ \sum_{p_1,\dots,p_r\in V(\GPo)}^{\not = }  
		\1_{\{D_{\GPo}(p_1)=\dots=D_{\GPo}(p_r)=k_n\}}\right] \\
	&=\frac{1}{r!} \int_\Rcal \cdots \int_\Rcal \varphi_{\Po}(\{p_1,\dots,p_r\},p_1,\dots,p_r)
		f(x_1,y_1) \dd x_1 \dd y_1 \cdots f(x_{r},y_{r})\dd x_{r} \dd y_{r},
\end{align*}
where we integrate over $r$ additional points which we can think of as being added independently and with the same distribution as the vertices of the poissonized KPKVB model $\GPo$ in the upper half-plane coordinates.

With $r=1$, it follows that
\[
	\Exp{N_{\Po}(k_n)} = \int_\Rcal \phi_{\Po}(\{p_1\}, k_n ; p_1) f(x_1,y_1) \dd x_1 \dd y_1,
\]
which yields that the right-hand side of the claim of the lemma can be rewritten as
\[
	\frac{1}{r!}(\Exp{N_{\Po}(k_n)})^r = \frac{1}{r!}\left(\int_\Rcal \phi_{\Po}(\{p_1\}, k_n; p_1) f(x_1,y_1) \dd x_1 \dd y_1,\right)^r.
\]

@@

Recall that, for any $0 < \varepsilon < 1$
\begin{align*}
	\Prob{D_{\Po} = k_n + t} &= \int_0^R \Prob{\Po(\mu_{Po,n}(y)) = k_n+ t} \alpha e^{-\alpha y} dy\\
	&= \int_0^{(1-\varepsilon)R} \Prob{\Po(\mu_{Po,n}(y)) = k_n+ t} \alpha e^{-\alpha y} dy \\
	&\hspace{10pt}+ \int_{(1-\varepsilon)R}^R \Prob{\Po(\mu_{Po,n}(y)) = k_n+ t} \alpha e^{-\alpha y} dy
\end{align*}
By Lemma~\ref{lem:degree_integral} the first part is $(1+o(1)) p_{k_n + t}$ while the second part is $O(n^{-2\alpha(1-\varepsilon)})$ and hence
\[
	\Prob{D_{\Po} = k_n + t} \le \bigO{1}\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right).
\]
In addition we have that $\Prob{Z(n,t) = t} \le \bigO{1} \frac{\Exp{Z(n,t)}}{t}$. We thus obtain
\begin{align*}
	\sum_{t = 1}^{N-n} \Prob{Z(n,t) = t}\Prob{D_{\Po} = k_n + t}
	&\le \bigO{1} \sum_{t = 1}^{N-n} \frac{\Exp{Z(n,t)}}{t}\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right)\\
	&= \bigO{1} \sum_{t = 1}^{N-n} \frac{(N-m)(k_n + t)}{N}\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right)\\
	&= \bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n} \frac{k_n}{t}
		\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right)\\
	&\hspace{10pt}+ \bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n}
		\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right),
\end{align*}
where we used that $\frac{N-n}{N} = \bigO{\sqrt{\frac{\log n}{n}}}$.

We will show that both summations are $\smallO{p_{k_n}}$. For the first summation we recall that $\frac{k_n (\log n)^{3/2}}{\sqrt{n}} = \smallO{1}$ for $k_n = \bigO{n^{\frac{1}{2\alpha + 1}}}$, while for $\varepsilon > 0$ small enough $n^{-2\alpha(1-\varepsilon)} = \smallO{k_n^{-(2\alpha + 1)}} = \smallO{p_{k_n}}$. Hence, since $p_{k_n + t} \le p_{k_n}$,
\begin{align*}
	\bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n} \frac{k_n}{t}
		\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right)
	&\le \bigO{ k_n \sqrt{\frac{\log n}{n}} p_{k_n}} \sum_{t = 1}^{N - m} \frac{1}{t}\\
	&= \bigO{\frac{k_n (\log n)^{3/2}}{\sqrt{n}}} p_{k_n} = \smallO{p_{k_n}}.
\end{align*}

For the other summation we use that
\[
	\sum_{t = 1}^{N-n} p_{k_n + t} \le \sum_{t = 1}^\infty p_{k_n + t}
	\le \bigO{k_n^{-2\alpha}} = \bigO{k_n p_{k_n}},
\]
together with the fact that for $\varepsilon$ small enough, $\log(n) n^{-2\alpha(1-\varepsilon)} = \smallO{k_n^{-(2\alpha + 1)}} = \smallO{p_{k_n}}$. This implies that
\begin{align*}
	\bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n}
		\left(p_{k_n + t} + n^{-2\alpha(1-\varepsilon)}\right)
	&\le \bigO{\sqrt{\frac{\log n}{n}} k_n p_{k_n}} + \bigO{(N-n)\sqrt{\frac{\log n}{n}} n^{-2\alpha(1-\varepsilon)}}\\
	&\le \smallO{p_{k_n}} + \bigO{(\log n) n^{-2\alpha(1-\varepsilon)}} = \smallO{p_{k_n}}.
\end{align*}

@@

We now have all necessary ingredient to proof the main result on the degrees, Theorem~\ref{thm:degrees_hyperbolic}.

>>>>

\section{Degrees when $k\to\infty$: proof of Theorem~\ref{thm:degrees_hyperbolic}\label{sec:degrees}}

@@

We start by using the Campbell-Mecke formula to compare the degree distribution in $\GPo$ with that of the typical point in $\Ginf$. 
As we've already seen this equals 
\[
	\pmf(k) = \int_0^\infty \rho(y,k) \alpha e^{-\alpha y} \, dy.
\]
We will relate this to the Poissonized KPKVB model $\GPo$. 
More precisely, let $N_{\Po}(k)$ denote the set of degree $k$ vertices in $\GPo$. 
We then show in Section~\ref{ssec:expected_degrees_GPo} that for any $1 \ll k_n \le n -1$ with 
$k_n = \bigO{n^{\frac{1}{2\alpha + 1}}}$, $\Exp{N_{\Po}(k_n)} = (1+\smallO{1}) n \pmf(k_n)$ and more generally,

\begin{equation}\label{eq:def_factorial_moments_GPo}
	\Exp{\binom{N_{\Po}(k_n)}{r}} = (1+\smallO{1}) \frac{\Exp{N_{\Po}(k_n)}^r}{r!},
\end{equation}

\noindent
for any integer $r \ge 1$ in Section~\ref{ssec:factorial_moments_GPo}. 
The latter result requires us to analyze the joint degree distribution in $\GPo$, which we do in Section~\ref{ssec:joint_degrees_GPo}. 
The above result in particular implies concentration of $N_{\Po}(k_n)$ from which the result on the degree distribution in $\GPo$ follows for $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. When $k_n = (1+\smallO{1}) c n^{\frac{1}{2\alpha + 1}}$ we use the above result to show that the fraction of degree $k_n$ nodes in $\GPo$ converges to a Poisson distribution.

@@

We will also establish all the above mentioned results for the finite box model $\Gbox$, since the proofs only require small 
alterations and we will need these results later on when analyzing the clustering coefficient and function.

@@

We can now use a concentration of heights argument to show that the integration of the Poisson probabilities $\Prob{\Po(\mu_{\Po}(y)) = k_n}$ over $0 \le y \le (1-\varepsilon) R$ is asymptotically equivalent to $\pmf(k_n)$. And the same holds if we instead consider $\mu_{\mathrm{box}}(y)$. The proof contains some technical elements that are contained in the Appendix to not hinder the flow of the argument.

\begin{lemma}\label{lem:degree_integral}
Let $0 < \varepsilon < 1$. Then for all $0 \le k_n \le n - 1$, as $n \to \infty$,

\begin{equation}\label{eq:degree_integral_BallHyp}
	\int_0^{(1-\varepsilon)R} \Prob{\Po(\Mu{\BallHyp{y}} = k_n} \alpha e^{-\alpha y} \dd y
	= (1+\smallO{1}) \pmf(k_n).
\end{equation}

Moreover, the same holds if we replace $\Mu{\BallHyp{y}}$ with $\Mu{\BallPon{y}}$ in~\eqref{eq:degree_integral_BallHyp}.
\end{lemma}

\begin{proof}
We will show that
\[
	\int_0^{(1-\varepsilon)R} \Prob{\Po(\mu_{\Po}(y)) = k_n} \alpha e^{-\alpha y} \dd y
	= (1+o(1))\int_0^\infty \Pee(\Po(\xi e^{\frac{z}{2}})=k_n) \alpha e^{-\alpha z} \dd z.
\]
This implies the result because the last integral equals $(1+o(1))2\alpha \xi^{2\alpha} k_n^{-(2\alpha+1)} = (1+o(1))\pmf(k_n)$.

@@

\begin{lemma}[First moment of number of degree $k$ vertices]\label{lem:expnnkn}
Let $N_{Po}(k)$ and $\Nbox(k)$ denote the number of vertices with degree $k$ vertices in the poissonized KPKVB model $\GPo$ and the finite box model $\Gbox$, respectively. Consider a sequence of integers $k_n\rightarrow\infty$ with $0 \leq k_n \leq n-1$.	If $k_n = O\left(n^{\frac{1}{2\alpha+1}}\right)$, then as $n \to \infty$,
\[
	\Exp{N_{Po}(k_n)} =(1+o(1)) n \pmf(k_n) \quad \text{and} \quad \Exp{\Nbox(k_n)} =(1+o(1)) n \pmf(k_n).
\]
Moreover, if $k_n \gg n^{\frac{1}{2\alpha + 1}}$ then
\[
	\Exp{N_{\Po}(k_n)} = \smallO{1} \quad \text{and} \quad \Exp{\Nbox(k_n)} = \smallO{1}.
\]
\end{lemma}

@@

When $k_n \gg n^{\frac{1}{2\alpha + 1}}$, Lemma~\ref{lem:degree_integral} implies that
\[
	n \int_0^{(1-\varepsilon)R} \Pee(Po(\mu_{Po}(y))=k_n) \alpha e^{-\alpha y}dy
	= (1+\smallO{1}) n \pmf(k_n)= \bigO{n k^{-(2\alpha + 1}} = \smallO{1}.
\]
On the other hand,
\[
	n \int_{(1-\varepsilon)R}^R \Pee(Po(\mu_{Po}(y))=k_n) \alpha e^{-\alpha y}dy = \bigO{n^{1 - 2\alpha(1-\varepsilon)}}.
\]
which is $\smallO{1}$ since by our choice $2\alpha(1-\varepsilon) > 1$. Thus the second claim of the lemma follows.
\end{proof}

@@

\begin{lemma}[Asymptotic factorization of degree probabilities]\label{lem:probdegFact}
Let $(k_n)$ be a sequence of integers with $0\leq k_n \leq n-1$, $k_n = \bigO{n^{\frac{1}{2\alpha+1}}}$ and $k_n \rightarrow \infty$.
Let $C, C^\prime>0$ and $r,s$ be positive integers with $r+1\leq s$. Fix $0<\varepsilon<1$. Then, it holds uniformly for all $(v_1,\dots,v_s) \in \left(\stripknc\right)^{s} = \stripknc \times \dots \times \stripknc$, satisfying $|x_{v_i}-x_{v_{r+1}}|\geq k_n^{1+\varepsilon}$ for all  $1\leq i \leq r$,  that
\[
	\varphi(\{v_1,\dots,v_{r+1}\};v_1,\dots,v_s)
	=(1+o(1))\varphi(\{v_1,\dots,v_r\};v_1,\dots,v_s)\varphi(\{v_{r+1}\};v_1,\dots,v_s)+\bigO{k_n^{-C^\prime}},
\]
where $\varphi$ is either $\varphi_{\Po}$ or $\varphi_{\mathrm{box}}$.
(Here uniformity means that the $\smallO{1}$ and $\bigO{k_n}$ terms do not depend on $v_1,\dots,v_{s}$.)
\end{lemma}

@@

We will now use that $\frac{(a+b)!}{a!} = (1+o(1))(a+b)^b$ for $b^2=o(a)$, applied to $a=k_n-y_{r+1}$ and $b=y_{r+1}-s$. To see this auxiliary fact, note that by Stirling's approximation to the factorial (see e.g.~\cite{Dutkay2013},~\cite{Nanjundiah1959}), it follows that

@@

\begin{lemma}[Factorial moments]\label{lem:factmoment}
Recall that $N_{Po}(k)$ denotes the number of degree $k$ vertices in $\GPo$. Let $(k_n)$ be a sequence of integers with $0\leq k_n \leq n-1$, $k_n = \bigO{n^{\frac{1}{2\alpha+1}}}$ and $k_n \rightarrow \infty$. Then, for any positive integer $r$, it holds that
\[
	\Exp{{N_{Po}(k_n) \choose r}} =(1+o(1)) \frac{(\Exp{ N_{Po}(k_n)})^r}{r!}.
\]
\end{lemma}

@@

\begin{lemma}[Factorization of degrees]
	\label{lem:asympind}
	Let $(k_n)$ be a sequence of integers with $0\leq k_n \leq n-1$, $k_n = \bigO{n^{\frac{1}{2\alpha+1}}}$ and $k_n \rightarrow \infty$.

	Let $\varphi$ be either $\varphi_{\mathrm{box}}$ or $\varphi_{\Po}$. Then we have that
	\begin{align*}
	&\int_\Rcal \cdots \int_\Rcal \varphi(\{p_1,\dots,p_r \}, k_n ;p_1,\dots,p_r) 
		f(x_1,y_1) \dd x_1 \dd y_1 \cdots f(x_r,y_r)\dd x_r \dd y_r \\
	&= (1+o(1))\left(\int_\Rcal  \varphi(\{p_1\}, k_n ;p_1) f(x,y) \dd x \dd y\right)^r .
	\end{align*}
\end{lemma}

@@

First of all, we observe that
\begin{align*}
{N_{\Po}(k_n) \choose r} = \frac{1}{r!} \sum_{p_1,\dots,p_r\in V(\GPo), \atop \text{distinct}} \1_{\{D_{\GPo}(p_1)=\dots=D_{\GPo}(p_r)=k_n\}}.
\end{align*}
This can be seen by induction on $r$. For $r=1$, the claim is clear. Assuming it holds for $r\geq 1$, by the induction hypothesis,
\begin{align*}
&{N_{\Po}(k_n) \choose r+1} = {N_{\Po}(k_n) \choose r} \frac{N_{\Po}(k_n)-r}{r+1} \\
&= \frac{1}{(r+1)!}\sum_{p_1,\dots,p_r\in V(\GPo), \atop \text{distinct}} \1_{\{D_{\GPo}(p_1)=\dots=D_{\GPo}(p_r)=k_n\}} (N_{\Po}(k_n)-r).
\end{align*}
Now, we can write
\[
	N_{\Po}(k_n) = \sum_{\stackrel{p_{r+1} \in V(\GPo),}{p_{r+1} \not \in\{ p_1,\dots,p_r\}} } \1_{\{D_{\GPo}(p_{r+1})=k_n\}}+\sum_{p_{r+1} \in \{p_1,\dots,p_r\}}\1_{\{D_{\GPo}(p_{r+1})=k_n\}}.
\]
The first sum leads to the right-hand side of the claim for $r+1$, whereas the second sum will cancel with the $-r$.

By the Campbell-Mecke formula 
\begin{align*}
	&\Exp{ {N_{\Po}(k_n) \choose r}}=\frac{1}{r!} \E\left[ \sum_{p_1,\dots,p_r\in V(\GPo), \atop \text{distinct}}  
		\1_{\{D_{\GPo}(p_1)=\dots=D_{\GPo}(p_r)=k_n\}}\right] \\
	&=\frac{1}{r!} \int_\Rcal \cdots \int_\Rcal \varphi_{\Po}(\{p_1,\dots,p_r\},p_1,\dots,p_r)
		f(x_1,y_1) \dd x_1 \dd y_1 \cdots f(x_{r},y_{r})\dd x_{r} \dd y_{r},
\end{align*}
where we integrate over $r$ additional points which we can think of as being added independently and with the same distribution as the vertices of the Poissonized KPKVB model $\GPo$ in the upper half-plane coordinates.

With $r=1$, it follows that
\[
	\Exp{N_{\Po}(k_n)} = \int_\Rcal \varphi_{\Po}(\{p_1\}, k_n ; p_1) f(x_1,y_1) \dd x_1 \dd y_1,
\]
which yields that the right-hand side of the claim of the lemma can be rewritten as
\[
	\frac{1}{r!}(\Exp{N_{\Po}(k_n)})^r = \frac{1}{r!}\left(\int_\Rcal \phi_{\Po}(\{p_1\}, k_n; p_1) f(x_1,y_1) \dd x_1 \dd y_1,\right)^r.
\]

@@

Recall that, for any $0 < \varepsilon < 1$
\begin{align*}
	\Prob{D_{\Po} = k_n + t} &= \int_0^R \Prob{\Po(\mu_{Po,n}(y)) = k_n+ t} \alpha e^{-\alpha y} dy\\
	&= \int_0^{(1-\varepsilon)R} \Prob{\Po(\mu_{Po,n}(y)) = k_n+ t} \alpha e^{-\alpha y} dy \\
	&\hspace{10pt}+ \int_{(1-\varepsilon)R}^R \Prob{\Po(\mu_{Po,n}(y)) = k_n+ t} \alpha e^{-\alpha y} dy
\end{align*}
By Lemma~\ref{lem:degree_integral} the first part is $(1+o(1)) \pmf(k_n + t)$ while the second part is $O(n^{-2\alpha(1-\varepsilon)})$ and hence
\[
	\Prob{D_{\Po} = k_n + t} \le \bigO{1}\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right).
\]
In addition we have that $\Prob{Z(n,t) = t} \le \bigO{1} \frac{\Exp{Z(n,t)}}{t}$. We thus obtain
\begin{align*}
	\sum_{t = 1}^{N-n} \Prob{Z(n,t) = t}\Prob{D_{\Po} = k_n + t}
	&\le \bigO{1} \sum_{t = 1}^{N-n} \frac{\Exp{Z(n,t)}}{t}\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right)\\
	&= \bigO{1} \sum_{t = 1}^{N-n} \frac{(N-m)(k_n + t)}{N}\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right)\\
	&= \bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n} \frac{k_n}{t}
		\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right)\\
	&\hspace{10pt}+ \bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n}
		\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right),
\end{align*}
where we used that $\frac{N-n}{N} = \bigO{\sqrt{\frac{\log n}{n}}}$.

We will show that both summations are $\smallO{\pmf(k_n)}$. For the first summation we recall that $\frac{k_n (\log n)^{3/2}}{\sqrt{n}} = \smallO{1}$ for $k_n = \bigO{n^{\frac{1}{2\alpha + 1}}}$, while for $\varepsilon > 0$ small enough $n^{-2\alpha(1-\varepsilon)} = \smallO{k_n^{-(2\alpha + 1)}} = \smallO{\pmf(k_n)}$. Hence, since $\pmf(k_n + t)\le \pmf(k_n)$,
\begin{align*}
	\bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n} \frac{k_n}{t}
		\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right)
	&\le \bigO{ k_n \sqrt{\frac{\log n}{n}} \pmf(k_n)} \sum_{t = 1}^{N - m} \frac{1}{t}\\
	&= \bigO{\frac{k_n (\log n)^{3/2}}{\sqrt{n}}} \pmf(k_n) = \smallO{\pmf(k_n)}.
\end{align*}

For the other summation we use that
\[
	\sum_{t = 1}^{N-n} \pmf(k_n + t) \le \sum_{t = 1}^\infty \pmf(k_n + t)
	\le \bigO{k_n^{-2\alpha}} = \bigO{k_n \pmf(k_n)},
\]
together with the fact that for $\varepsilon$ small enough, $\log(n) n^{-2\alpha(1-\varepsilon)} = \smallO{k_n^{-(2\alpha + 1)}} = \smallO{\pmf(k_n)}$. This implies that
\begin{align*}
	\bigO{\sqrt{\frac{\log n}{n}}} \sum_{t = 1}^{N-n}
		\left(\pmf(k_n + t) + n^{-2\alpha(1-\varepsilon)}\right)
	&\le \bigO{\sqrt{\frac{\log n}{n}} k_n \pmf(k_n)} + \bigO{(N-n)\sqrt{\frac{\log n}{n}} n^{-2\alpha(1-\varepsilon)}}\\
	&\le \smallO{\pmf(k_n)} + \bigO{(\log n) n^{-2\alpha(1-\varepsilon)}} = \smallO{\pmf(k_n)}.
\end{align*}

@@

We now have all necessary ingredients to prove the main result on the degrees, Theorem~\ref{thm:degrees_hyperbolic}.
